---
title: "PCA on Diet Classification Data"
output: html_notebook
---

```{r}
library(tidyverse)
library(cluster)
library(ggplot2)
```


```{r}
diet_data <- classified_df
```

```{r}

numeric_vars_scaled <- diet_data %>%
  select(where(is.numeric)) %>%
  select_if(~ var(.) != 0) %>%  # Removes columns with zero variance
  scale()

# Run PCA on the scaled data
pca_result <- prcomp(numeric_vars_scaled, center = TRUE, scale. = TRUE)

# UPGMA clustering on PCA scores
distances <- dist(pca_result$x, method = "euclidean")
upgma_cluster <- hclust(distances, method = "average")

# Plot UPGMA dendrogram
plot(upgma_cluster, main = "UPGMA Dendrogram based on PCA Scores", xlab = "Species", sub = "", cex = 0.7)

# Summary of explained variance for the first four components
variance_explained <- summary(pca_result)$importance[2, 1:4]
print(variance_explained)

# Prepare PCA data for plotting with Diet_Class color-coded grouping
pca_df <- as.data.frame(pca_result$x)
pca_df$Diet_Class <- diet_data$Diet_Class  # Adding Diet_Class for grouping

# Define color palette for Diet_Class groups
color_palette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
                   "#0072B2", "#D55E00", "#CC79A7", "#000000", "maroon")

# Plot PCA scatter plot with color-coded grouping by Diet_Class
ggplot(pca_df, aes(x = PC1, y = PC2, color = Diet_Class)) +
  geom_point(size = 2) +
  labs(title = "PCA Plot with Z-scored Variables", x = "PC1", y = "PC2") +
  scale_color_manual(values = color_palette) +
  theme_minimal() +
  theme(legend.title = element_text(size = 10), legend.text = element_text(size = 8))

# Save the PCA plot as PDF (optional)
ggsave("C:/Users/umder/Desktop/meyerlab/Data/pca_plot_diet_class.pdf", width = 8, height = 6)